\begin{publication}

\subsubsection*{Peer-Review Journals Papers}

\begin{enumerate}\scriptsize
	\item \textcolor{red}{\textbf{N. Piasco}, D. Sidib\'e, V. Gouet-Brunet and C. Demonceaux. Auxiliary Modality as Robust Training Signal for
Visual Localization in Challenging Conditions.}
	\item \textbf{N. Piasco}, D. Sidib\'e, C. Demonceaux and V. Gouet-Brunet. A Survey on Visual-Based Localization: On the Benefit of Heterogeneous Data. Pattern Recognition, Volume 74, February 2018, pp.90-109.
\end{enumerate}

\subsubsection*{Peer-Review International Conferences}

\begin{enumerate}\scriptsize
	\item \textbf{N. Piasco}, D. Sidib\'e, C. Demonceaux and V. Gouet-Brunet. Perspective-n-Learned-Point: Relative Pose
Estimation at Absolute Scale. 2019 British Machine Vision Conference (BMVC), Cardiff, United Kingdom, September 2019.
	\item \textbf{N. Piasco}, D. Sidib\'e, C. Demonceaux and V. Gouet-Brunet. Geometric Camera Pose Refinement with Learned Depth Maps. 2019 IEEE International Conference on Image Processing (ICIP), Taipei, Taiwan, September 2019.
	\item \textbf{N. Piasco}, D. Sidib\'e, V. Gouet-Brunet and C. Demonceaux. Learning Scene Geometry for Visual Localization in Challenging Conditions. 2019 IEEE International Conference of Robotics and Automation (ICRA), Montreal, Canada, May 2019.	
\end{enumerate}

\subsubsection*{Peer-Review National Conferences}

\begin{enumerate}\scriptsize
	\item \textbf{N. Piasco}, D. Sidib\'e, V. Gouet-Brunet and C. Demonceaux. Apprentissage de modalit\'es auxiliaires pour la localisation bas\'ee vision. Reconnaissance des Formes, Image, Apprentissage et Perception (RFIAP), Champs-sur-Marne, France, June 2018
	\item \textbf{N. Piasco}, D. Sidib\'e, V. Gouet-Brunet and C. Demonceaux. Localisation Bas\'ee Vision : de l’h\'et\'erog\'en\'eit\'e des approches et des données. ORASIS - Journ\'ees francophones des jeunes chercheurs en vision par ordinateur (ORASIS), Colleville-sur-Mer, France, June 2017.
\end{enumerate}

\subsubsection*{Thesis}

\begin{enumerate}\scriptsize
\item \textbf{N. Piasco}. Vision-based localization with discriminative features from heterogeneous visual data.
\end{enumerate}

\end{publication}
